---
title: "QBS 181 HW3"
author: "Natt Chan"
date: "11/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
install.packages("dplyr", repos="http://cran.us.r-project.org")
library(dplyr)
library(tidyverse)
library(magrittr)
install.packages("nycflights13", repos="http://cran.us.r-project.org")
library(nycflights13)
library(lubridate)
```

\section{1. Compute the rate for table2, and table4a+table4b and perform the four operation}

check out dplyr cheat sheet

```{r 1a}
table2.cases <- filter(table2, type == "cases") %>%
  rename(cases = count) %>%
  arrange(country, year)

table2.cases

```

```{r 1b}
table2.pop <- filter(table2, type == "population") %>%
  rename(population = count) %>%
  arrange(country, year)

table2.pop
```
```{r 1c}
table2.rate <- tibble(
  year = table2.cases$year,
  country = table2.cases$country,
  cases = table2.cases$cases,
  population = table2.pop$population
) %>%
  mutate(rate = (cases / population) * 10000) %>%
  select(country, year, rate)

table2.rate
```
```{r 1d}
table2.rate <- table2.rate %>%
  mutate(type = "rate") %>%
  rename(count = rate) %>%
  select(country, year, type, count)
table2.rate

(bind_rows(table2, table2.rate) %>%
  arrange(country, year, type, count))

```
For table4a and table4b:

```{r 1e}
new.table4 <- tibble(
    country = table4a$country,
    `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
    `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000
  )
new.table4

```

\section{2. Why does this code fail?}

```{r 2}
#table4a%>%gather(1999,2000,key="year",value="cases")
```

It is because we are trying to work with columns that do not exist.

\section{3. Use the flights dataset in the nycflights13 library and answer the following:}


\subsection{a. How does the distribution of flights times within a day change over the course of the year?}

```{r 3a}
flights
#avg air time each day over course of year
#sep column to consolidate times
#summarise mean air time after grouping by day

arrange(flights,month,year,day)
avg_flight_time <- flights %>% filter(!is.na(air_time)) %>% group_by(year,month,day) %>% summarise(mean(air_time))
head(avg_flight_time)

```

\subsection{b. Compare dep_time,sched_dep_time, and dep_delay. Are they consistent. Explain your findings:}

```{r 3b}
comparisons <- flights_jan1 %>%
  select(dep_time, sched_dep_time, dep_delay)
head(comparisons)
```

\subsection{c. Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint:create a binary variable that tells whether or not a flight was delayed.}
